### 2017-08-27 src/4_OnlineLearning.tex
explicacion de ridge regression
last comments chapter 3

%------------------------FIN DRAFT PROFE -------------------------------
{\em Supervised learning\/} consists then in finding a {\em learning function\/}
$f:X\to Y$ that minimisr the expected error of the loss function $V(f(\mathbf{x}),y): Y
\times Y$ defined as:

\begin{eqnarray*}
\label{eq:lossfunction}
 V & \rightarrow & \mathbb{R}^+ \\
  (f(\mathbf{x}), y) &\rightarrow & V(f(\mathbf{x}),y)
\end{eqnarray*}

@@

Since the true distribution is unknown and only training samples are available,
the objective is to estimate a function $\hat{f}$ through empirical risk
(training error) minimisation (ERM):

\begin{equation*}
\hat{f} = \argmin_{f \in \mathscr{H}} R_{\text{emp}}[f] 
\end{equation*}

@@

\begin{equation}
\frac{1}{n} \sum_{i=1}^n V(f(x_i),y_i) \xrightarrow{n \rightarrow \infty}  E[V(f(\mathbf{x}),y)]
\end{equation}

@@

\begin{figure}[!h]
  \centering
  \includegraphics[width=0.8\textwidth]{img/VCdimension}
  \caption{VC dimension example from \cite{vapnik2013nature}. The left figure
  shows that three vectors can be shattered by a line in a plane. However, if we
  add an extra vector $z_4$, it is impossible to shatter them with a
  plane.Therefore, linear classifiers has VC dimension equals to $d+1$ where $d$
  is the dimension of the instance space $X$.}
  \label{fig:vcdimension}
\end{figure}

@@

Ridge regression (RR) was independently proposed by Tikhonov \cite{tikhonov1963}
and Phillips \cite{phillips1962}. RR corresponds to a regularised least squares
method. The least squares (LS) method is a well known way to solve a regression
problem. LS method consists of minimising the sum of squared errors:

\begin{eqnarray*}
\label{eq:lsm}
 J(\mathbf{X}) =& \displaystyle \sum_{t=1}^N (f(\mathbf{x}_t)-y_t)^2 \\
 =&\displaystyle \sum_{t=1}^N (\mathbf{\mathbf{A}}^\top {\mathbf{x}}_t-y_t)^2 \\
 =& \| \mathbf{A}\mathbf{\mathbf{X}} - \mathbf{Y} \|_2^2 
\end{eqnarray*}

>>>>

%------------------------FIN DRAFT PROFE -------------------------------
{\em Supervised learning\/} consists then in finding a {\em learning function\/}
$f:X\to Y$ that minimise the expected error of the loss function $V(f(\mathbf{x}),y): Y
\times Y$ defined as:

%\begin{eqnarray*}
%\label{eq:lossfunction}
% V & \rightarrow & \mathbb{R}^+ \\
%  (f(\mathbf{x}), y) &\rightarrow & V(f(\mathbf{x}),y)
%\end{eqnarray*}

\begin{equation}\label{eq:lossfunction}
V(f(\cdot),\cdot):X\times Y\to\R_0^+\,,\qquad
(\bfx,y)\mapsto V\big(f(\bfx),y\big)
\end{equation}

@@

Since the true distribution is unknown and only training samples are available,
the objective is to estimate a function $\widehat{f}$ through empirical risk
(training error) minimisation (ERM):

\begin{equation*}
\widehat{f} = \argmin_{f \in \mathscr{H}} R_{\text{emp}}[f] 
\end{equation*}

@@

\begin{equation}
%\frac{1}{n} \sum_{i=1}^n V(f(x_i),y_i) \xrightarrow{n \rightarrow \infty}  E[V(f(\mathbf{x}),y)]
\frac1n\,\sum_{i=1}^n V\big( f(\bfx_i,y_i \big)
\overset{n\to\infty}{\longrightarrow} E\big[ V(f(\cdot),\cdot)\big]\,.
\end{equation}

@@

\begin{figure}[!h]
  \centering
  \includegraphics[width=0.8\textwidth]{img/VCdimension}
  \caption{VC dimension example from \cite{vapnik2013nature}. The left figure
  shows that the 3 dichotomies can be shattered by a line in a plane. However, if we
  add an extra vector $z_4$, it is impossible to shatter them with a
  line. Therefore, linear classifiers has VC dimension equals to $d+1$ where $d$
  is the dimension of the instance space $X$.}
  \label{fig:vcdimension}
\end{figure}

A key point is Radon's theorem in discrete geometry \cite{peterson1972geometry}
($\text{co}(A)$ denotes the convex hull of $A$):

\begin{theorem}[J. Radon]
Let $S\subset\R^d$ be a subset containing at least $d+2$ elements. 
Then there exist {\bf disjoint} subsets $A\subset S$ and $B\subset S$
such that
$$
\text{co}(A)\cap\text{co}(B)\neq\emptyset
$$
\end{theorem}

Radon's theorem implies that in $\R^d$ any set of $d+2$ or more vectors has at least one dichotomy which cannot be
separated (shattered) by any plane in $\R^d$. Thus the VC-dimension in $\R^d$ cannot be greater than $d+1$.

@@

Ridge regression (RR) was independently proposed by Tikhonov \cite{tikhonov1963}
and Phillips \cite{phillips1962}.

RR corresponds to a regularised least squares
method. The least squares (LS) method is a well known way to solve a regression
problem.

\begin{equation}
\label{eq:RRdefinitions}
\mathbf{X} = 
\left[
  \begin{array}{cccc}
    \vertbar & \vertbar &        & \vertbar \\
    \mathbf{x}_{1}    & \mathbf{x}_{2}    & \ldots & \mathbf{x}_{n}    \\
    \vertbar & \vertbar &        & \vertbar 
  \end{array}
\right]
\quad , \quad
\mathbf{A} = 
\left[
  \begin{tabular}{c>{$}c<{$}c}
    --- & \mathbf{a}^{\top}_{1} & ---\\
    --- & \mathbf{a}^{\top}_{2} & ---\\
    & \vdots & \\
    --- & \mathbf{a}^{\top}_{m} & ---
  \end{tabular}
\right]
\quad \text{and} \quad
\mathbf{Y} = 
\left[
  \begin{array}{cccc}
    \vertbar & \vertbar &        & \vertbar \\
    \mathbf{y}_{1}    & \mathbf{y}_{2}    & \ldots & \mathbf{y}_{n}    \\
    \vertbar & \vertbar &        & \vertbar 
  \end{array}
\right]
\end{equation}

LS method consists of minimising the sum of squared errors:
\begin{eqnarray*}
\label{eq:lsm}
 J(\mathbf{X}) =& \displaystyle \sum_{t=1}^n (f(\mathbf{x}_t)-y_t)^2 \\
 =&\displaystyle \sum_{t=1}^n (\mathbf{\mathbf{A}}^\top {\mathbf{x}}_t-y_t)^2 \\
 =& \| \mathbf{A}\mathbf{\mathbf{X}} - \mathbf{Y} \|_2^2 
\end{eqnarray*}
